Supplementary Materials

### 2022-12-07 manuscript/manuscript.tex
supp draft

@@

% \section*{Introduction}

@@

@@

\begin{align}
  \label{eq:q_t}
  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^T (1-r)^T\right)}{\kappa+(1-\kappa)r}
\end{align}

which as $T \to \infty$, converges to,

@@

\begin{align}
      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{\infty,l}^2\right) \\
              &\approx N \exp\left(- \frac{V_{h}}{L} \int_0^{\nicefrac{L}{2}} Q_{\infty}(r)^2 dr \right) \\
\end{align}

where $L = n r_\text{BP}$ is the total segment length in Morgans. Then, the
total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,

\begin{align}
  Q_\infty^2 &= \frac{2}{L}\int_0^{\nicefrac{L}{2}} Q_{\infty}(r)^2 dr  \\
  &=  \frac{2}{L} \int_0^{\nicefrac{L}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
  &= \frac{2}{(1-Z) (2-(2-L) Z)}
\end{align}

@@

\begin{align}
  Q_\infty^2 = \frac{2}{L (1-Z)} + \frac{2 (L-2)}{L^2}+\frac{2 \left(L^2-4 L+4\right) (1-Z)}{L^3} + \mathcal{O}(\kappa^2)
\end{align}

Santiago and Caballero's approximation only keeps this first term, e.g. their
$Q_\infty^2 \approx 2/(L(1-Z))$, which is only accurate in the domain $L >
0.2$.

\begin{align}
  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)L}\right) 
\end{align}

(c.f. \cite{Santiago1998-bs} equation 8). For small $L$, the small $\kappa$
approximation strongly deviates from the full equation; thus, throughout, we
use the unapproximated version,

\begin{align}
  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-L) Z)} \right).
\end{align}

@@

\begin{align}
  \label{eq:net_integral}
  N_{e,t} \approx N \exp\left(-\frac{V_h}{L} \int_0^{\nicefrac{L}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
\end{align}

@@

\begin{align}
  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + L - Ls} \right)
\end{align}

@@

At equilibrium, the $N_e$ that determines fixation probability is the
asymptotic $N_{e,\infty}$. Consequently, Santiago and Caballero argue that the
asymptotic reduction in effective population size is given by the following
system of non-linear equations,

\begin{align}
  R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-L) Z)} \right) \\
  N_{e,\infty} &= N \exp\left(-\frac{(U-R(N_{e,\infty}))s}{\left(1-\frac{Us}{U-R(N_{e,\infty})}\right) \left(2-(2-L) \frac{Us}{U-R(N_{e,\infty})}\right)} \right).
\end{align}

An analogous system can be found for $N_{e,t}$ using equation
\eqref{eq:net_integral}.

>>>>

% line numbers:
\RequirePackage{lineno}
%\modulolinenumbers[5]
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}

@@

\section*{Introduction}
\section*{Results}

\begin{figure}[!htb]
  \centering
  \includegraphics{figures/experimental_1.pdf}
  \caption{}
  \label{fig:figure-1}
\end{figure}

@@

TODO we need to give some intuition for what Vm and V are in humans, e.g. back of envelope type stuff

@@

\begin{align}
  \label{eq:q_t}
  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
\end{align}

which as $t \to \infty$, converges to,

@@

\begin{align}
      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
\end{align}

where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,

\begin{align}
  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \\
  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
  &= \frac{2}{(1-Z) (2-(2-M) Z)}
\end{align}

@@

\begin{align}
  Q_\infty^2 = \frac{2}{M (1-Z)} + \frac{2 (M-2)}{M^2}+\frac{2 \left(M^2-4 M+4\right) (1-Z)}{M^3} + \mathcal{O}(\kappa^2)
\end{align}

Santiago and Caballero's approximation only keeps this first term, e.g. their
$Q_\infty^2 \approx 2/(M(1-Z))$, which is only accurate in the domain $M >
0.2$.

\begin{align}
  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
\end{align}

(c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
approximation strongly deviates from the full equation; thus, throughout, we
use the unapproximated version,

\begin{align}
  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
\end{align}

@@

\begin{align}
  \label{eq:net_integral}
  N_{e,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
\end{align}

@@

\begin{align}
  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right)
\end{align}

@@

At equilibrium, the $N_e$ that determines fixation probability is the
asymptotic $N_{e,\infty}$. Consequently, Santiago and Caballero argue that the
asymptotic reduction in effective population size uses the $Q_\infty^2$ and is
given by the following system of non-linear equations,

\begin{align}
  R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-M) Z)} \right)
\end{align}

The solution to these equations is the asymptotic equilibrium $\widetilde{N}_e$ and
equilibrium ratchet rate $\widetilde{R}$. This set the equilibrium variance,
$\widetilde{V} = Us - 2\widetilde{R}s$.

\subsection*{Heterozygosity and $B$ values Under Weak and Strong Selection}

While the selection dynamics are set by the asymptotic $Q_\infty^2$ that
determines the equilibrium $\widetilde{N}_e$ and $\widetilde{R}$,
heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
term is

XXX

\subsection*{Numeric methods for calculating $B$}

Our $B$ is,

\begin{align}
  B = \frac{1}{2N} \sum_{t=0}^\infty \prod_{i=0}^t \left(1 - \frac{1}{2N_e(i)}\right).
\end{align}

This sum takes many terms to converge, which makes calculating the $B$ between
each focal neutral site and segment intractably slow (even when the equilibrium
$\widetilde{V}$ for the segment pre-calculated). We approximate this in the
following way for faster numerical calculations. Let $g(t) = \prod_{i=0}^t (1 -
\nicefrac{1}{2N_e(i)})$. Then, taking the first-order series expansion around
the logarithm of the terms and approximating the sum with an integral,

\begin{align}
  g(t) &\approx \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) 
       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
\end{align}

$B$ is the case of $\lim b \to \infty$ for $B(b)$. Approximating the sum in
$B(b)$ with an integral,

\begin{align}
  B(b) &= \int_0^b \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) dt \\
       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
\end{align}

Now, let $z = \log(t)$, then

\begin{align}
  B(w) &= \int_0^w \exp\left(- \frac{1}{2N} \int_0^{\log(z)} \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) \exp(t) dz \\
       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
\end{align}
